---
title: "R Presentations for the CERN Study Group"
author: "Achintya Rao"
date: "17 February 2017"
output: 
    revealjs::revealjs_presentation:
        theme: serif
        highlight: pygments
        incremental: true
        center: true
        transition: slide
        reveal_options:
            slideNumber: true
        fig_caption: true
        showNotes: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plotly)
```

---

<section><h1>Some eye-candy</h1>

## Responsive presentations with interactivity

- Example repo: [github.com/RaoOfPhysics/201604_PCST](https://github.com/RaoOfPhysics/201604_PCST)
- In action: [via.hypothes.is/https://raoofphysics.github.io/201604_PCST/](https://via.hypothes.is/https://raoofphysics.github.io/201604_PCST/)
- File generating the presentation: [raw.githubusercontent.com/RaoOfPhysics/201604_PCST/gh-pages/index.Rmd](https://raw.githubusercontent.com/RaoOfPhysics/201604_PCST/gh-pages/index.Rmd)
- Zenodo DOI: [10.5281/zenodo.50791](http://dx.doi.org/10.5281/zenodo.50791) 

</section>

---

<section><h1>HOWTO</h1>

## Tool chain

- [R](https://www.r-project.org/) + [RStudio](https://www.rstudio.com/)
- [R Markdown](http://rmarkdown.rstudio.com/) 
[[Cheatsheet PDF](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)] -- authoring
- [ggplot2](http://ggplot2.org/) -- making plots
- [plotly](https://plot.ly/) -- plot interactivity
- [reveal.js](http://lab.hakim.se/reveal-js/) -- displaying
- [GitHub Pages](https://pages.github.com/) -- sharing
- [Hypothesis](https://hypothes.is/) -- annotating

## Workflow

- Create a directory for your project
- Open RStudio
- Create a new `R Markdown` HTML file named `index.Rmd`
- Switch to console run:
    - `install.packages("ggplot2")`
    - `install.packages("plotly")`
    - `install.packages("revealjs")`

## Workflow (contd.)

- Add [YAML](http://www.yaml.org/) frontmatter!
- Instructions for reveal.js presentations: [rmarkdown.rstudio.com/revealjs_presentation_format.html](http://rmarkdown.rstudio.com/revealjs_presentation_format.html)
- Create sections and add content
- Knit your presentation!

</section>

---

<section><h1>Using R code in your presentation</h1>

- Source: [https://jofrhwld.github.io/avml2012/](https://jofrhwld.github.io/avml2012/)

## Example of some code

```{r cars}
summary(mtcars)
```

## Static plots

```{r, echo=FALSE}

# Code from https://jofrhwld.github.io/avml2012/
# Taken as an example only under fair-use practice
# All rights reserved by the original author

p <- ggplot(mtcars, aes(x = factor(1), fill = factor(cyl))) + 
  geom_bar(width = 0.8, position = "dodge", color = "black")

p
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent 
printing of the R code that generated the plot.

## Interactive plots!

```{r, echo=FALSE}

# Code from https://jofrhwld.github.io/avml2012/
# Taken as an example only under fair-use practice
# All rights reserved by the original author

ggplotly(p)
```

## Import some data

```{r}
I_jean <- read.delim("http://bit.ly/avml_ggplot2_data")
head(I_jean)
```

## Make more plots!

```{r, echo=FALSE}

# Code from https://jofrhwld.github.io/avml2012/
# Taken as an example only under fair-use practice
# All rights reserved by the original author

p <- ggplot(I_jean, aes(x = Dur_msec, y = F1.n)) + 
  geom_point() + 
  stat_smooth(method="loess") +
  scale_x_log10(breaks = c(50, 100,200,300,400)) +
  scale_y_reverse() +
  ylab("Normalized F1") +
  xlab("Vowel duration (msec)") +
  theme_bw() +
  labs(title = "394 tokens of 'I' from one speaker")

ggplotly(p)
```

</section>

---

<section><h1>Versioning, sharing, archiving</h1>

## Git going!

- `$ git init`
- Create a `.gitignore` and include `.RData` and `.Rhistory`
- Add `index.Rmd` and `index.html`
- Commit!

## Hosting your presentation on GitHub

- Create a GitHub repository to store your code and presentation
- Push to GitHub
- Activate GitHub Pages
- **Important!** Don't forget to include `README` and `LICENSE` files!
- This presentation: [github.com/RaoOfPhysics/201702_R-presentations](https://github.com/RaoOfPhysics/201702_R-presentations)
- Point to the presentation website in your repo's description…
- … with the [*via* proxy](https://hypothes.is/blog/meetvia/) from hypothesis 
for 
annotations

## Commit early, commit often

- Preparing your presentation is an opportunity for collaboration
- Send all your commits to GitHub and deploy your presentation incrementally
- Receive feedback in the form of issues or pull requests

## Archiving on Zenodo

- [Zenodo](https://zenodo.org/) is a service provided by CERN for archiving all 
kinds of research outputs, including presentations
- After giving your talk (with the final file), create a "Release" and send it 
off to Zenodo
    - [guides.github.com/activities/citable-code/](https://guides.github.com/activities/citable-code/)
    - [zenodo.org/account/settings/github/](https://zenodo.org/account/settings/github/)
- Mint a DOI and add a badge to your `README`

</section>

---

## Questions?

Discussion here: [github:CERNStudyGroup/cernstudygroup.github.io #46](https://github.com/CERNStudyGroup/cernstudygroup.github.io/issues/46)